---
title: "Convex combination of copulas"
author: "Vignette Author"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    fig_caption: yes
vignette: >
  %\VignetteIndexEntry{Convex combination of copulas}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

```

The goal of this vignette is to describe what are convex combinations of copulas and how we implemented them inside the `empCop` package. The actual implemntation only deals with *linear* combinations.

First, if we supose a list of copulas $C_i, i \in {1,...,n}$ all of the same dimension, we should emphasis that $\sum\limits_{i=1}^n \alpha_i C_i$ is still a copula as soon as $\sum\limits_{i=1}^n \alpha_i = 1$ and $\alpha_i >=0 \forall i$. 

This simple *linear* combinaison of copulas can be constructed in R through the following code : 

```{r}
library(copula)
library(empCop)

copulas <- list(
  archmCopula("gumbel",3,dim=2),
  archmCopula("clayton",-1,dim=2)
)
alpha <- c(1,4)

(cop <- ConvexCombCopula(copulas,alpha))

plot(rCopula(100,cop))
pCopula(c(0.5,0.7),cop)


```


The number of copulas is not restricte. Avaliables methods are for the moment `rCopula` and `pCopula`. 


