---
title: "Convex combination of checkerbords with known margins"
author: "Oskar Laverny"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    fig_caption: yes
vignette: >
  %\VignetteIndexEntry{Convex combination of checkerbords with known margins}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

```

The goal of this vignette is to describe what are convex combinations of copulas and how we implemented them inside the `empCop` package. The actual implemntation only deals with *linear* combinations.

First, if we supose a list of copulas $C_i, i \in {1,...,n}$ all of the same dimension, we should emphasis that $\sum\limits_{i=1}^n \alpha_i C_i$ is still a copula as soon as $\sum\limits_{i=1}^n \alpha_i = 1$ and $\alpha_i >=0 \forall i$. 

This simple *linear* combinaison of copulas can be constructed in R through the following code : 

```{r}
library(copula)
library(empCop)

copulas <- list(
  archmCopula("gumbel",3,dim=2),
  archmCopula("clayton",-1,dim=2)
)

(cop <- ConvexCombCopula(copulas))

plot(rCopula(100,cop))
pCopula(c(0.5,0.7),cop)


```


The number of copulas is not restricte. Avaliables methods are for the moment `rCopula` and `pCopula`. 


For exemple, given a dataset, we could optimise each copula first and them optimise weights, based on some criterion, e.g rmse. 


```{r}
data("LifeCycleSavings")
pseudo_dataset <- apply(LifeCycleSavings[sample(1:50,size=20),],2,rank)/(nrow(LifeCycleSavings)+1)
d <- ncol(pseudo_dataset)
pairs(pseudo_dataset)
```

For exemple, let's start from checkerboards(m=10) with known bivariate clayton margins. Since only the variables 2 to 4 seems to be non-independant, we will calculate cbkmCopulas only for thoose.

```{r}

cbkm_clayton <- function(data,margins,m=10){
  
  known_copula <- 
    archmCopula("Clayton",dim=2,
                param = coef(fitCopula(
                  archmCopula("Clayton",dim = length(margins)),
                  data[,margins],
                  method="itau")))
  
  cbkmCopula(data,
             m=m,
             pseudo=TRUE,
             margins_numbers = margins,
             known_cop = known_copula,
             quiet=TRUE
  )
}

couples <- as.matrix(expand.grid(2:4,2:4))
couples <- couples[couples[,1]<couples[,2],]

copulas <- list()
copulas[[1]] <- cbCopula(pseudo_dataset,pseudo=TRUE,m=4)

# for(i in 1:nrow(couples)){
#   copulas[[i+1]] <- cbkm_clayton(pseudo_dataset,couples[i,])
# }


```


For exemple, we can draw a simulations from the first cbkmCopula : 

```{r}
#pairs(rCopula(1000,copulas[[1]]))
```

Now that we have all thoose fits, we can optimise weights regarding a certain criterion. 

Here the criterion will be the RMSE on the dataset : 

```{r}
# rmse <- function(alpha,cop,data,pEmpCop){
#     sqrt(sum((pCopula(u=data,copula=cop) - pEmpCop)^2))
# }
# 
# criterion <- function(alpha = rep(1/length(cops),length(cops)),cops,data){
#   pEmpCop <- pCopula(u=data,copula=empCopula(data))
#   cop <- ConvexCombCopula(copulas = cops,alpha = alpha)
#   rmse(alpha,cop,data,pEmpCop)
# }

#criterion(cops = copulas,data = pseudo_dataset)
```


This implmeentation is not finished because the actual implementation of the pCopula function is clearly not good enough (not fast enough) to include cbkmCopulas in this process. 

The solution to spee things up is to pass coputations to C++ instead of R. Will do soon.

But for the moment this is not possible.








