---
title: "Empirical Checkerboard Copula with known margins"
author: "Vignette Author"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    fig_caption: yes
vignette: >
  %\VignetteIndexEntry{Empirical Checkerboard Copula with known margins}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
library(copula)
library(empCop)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 7
)
```


The empirical checkerboard copula with known margins is a model for copula that uses priori information on a multidimentional margins to condition a checkerboard construction on the rest of the copula. This vignette describes the model in the first section, and then discuss current implementation in the second section

## ECBC with known margins

### Preliminary notations


First of all, for a certain dimension of the data at hand $d$ and a certain checkerboard parameter $m$, let's consider the ensemble of multi-indexes $\mathcal{I} = \{\mathbf{i} = (i_1,..,i_d) \subset \{1,...,m \}^d\}$ which indexes the boxes : 

$$B_{\mathbf{i}} = \left]\frac{\mathbf{i}-1}{m},\frac{\mathbf{i}}{m}\right]$$

partitioning the space $\mathbb{I}^d = [0,1]^d$. Denote the set of thoose boxes by $\mathcal{B}_\mathcal{I} = \left\{B_{\mathbf{i}}, \mathbf{i} \in \mathcal{I}\right\}$. Furthermore, let's choose $p$ dimensions that would be assigned to a known copula, by setting : $J \subset \{1,...,d\}, |J| = p$ and let's define proper projections for the boxes : 

$$B^J_{\mathbf{i}} = \{ \mathbf{x} \in [0,1]^p, x_j \in \left]\frac{i_j-1}{m},\frac{i_j}{m}\right] \forall j \in J \}$$

$$B^{-J}_{\mathbf{i}} = \{\mathbf{x} \in [0,1]^p, x_j \in \left]\frac{i_j-1}{m},\frac{i_j}{m}\right] \forall j \notin J \}$$

such that $B_{\mathbf{i}} = B^J_{\mathbf{i}} \times B^{-J}_{\mathbf{i}}$ for all $\mathbf{i} \in \mathcal{I}$. The tensor product is here understood taking dimensions of $\mathbb{I}^d = [0,1]^d$ in the right order such that $B^J_{\mathbf{i}}$ dimensions end up in place of dimensions indexed by $J$. Think of dimensions as re-ordered such that $J = \{1,...,p\}$.

Let now $\lambda$ be the dimension-unspecified lebesgue measure on any power of $\mathbb{R}$, that is : 

$$\forall d \in \mathbb{N}, \forall \mathbf{x},\mathbf{y} \in \mathbb{R}^p, \lambda(\left[\mathbf{x},\mathbf{y}\right]) = \prod\limits_{p=1}^{d} (y_i - x_i)$$

Let furthermore $\mu^J$ be a copula measure of dimension $p$, corresponding to the known multivariate margin associated to marginals in $J$. Let also $\mu$ and $\hat{\mu}$ be dimensionaly-unspecific  version of the true copula measure of the sample at hand and (respectively) the classical Deheuvels empirical copula, that is : 

- For $n$ i.i.d obeservation of the copula of dimension $d$, let $\forall i \in \{1,...,d\}, \, R_i^1,...,R_i^d$ be the marginal ranks for the variable $i$. 
- $\forall \mathbf{x} \in \mathcal{I}^d, \text{ let } \hat{\mu}([0,x]) = \frac{1}{n} \sum\limits_{k=1}^n \mathbb{1}_{R_1^k\le x_1,...,R_d^k\le x_d}$


We are now ready to define the empirical checkerboard copula with known margins. 

### Definition, estimation and simulation procedures

The empirical copula with known margins is the copula that correspond to the following simulation procedure, which is the starting point of our reflexion. 

- Simulate a sample from the known sub-copula $\mu^J$, of dimension $p$, through any avaliable method (depends on the known copula model). Let $B_{\mathbf{i}}^J$ be the (projected) box containing this sample. 
- Sample one box $B_{\mathbf{i}}$ among all boxes with projection $B_{\mathbf{i}}^J$ with probability weights :
  -  $\frac{\hat{\mu}(B_{\mathbf{i}})}{\hat{\mu}(B_{\mathbf{i}}^J)}$ if the projected box $B_{\mathbf{i}}^J$ contains one or more (empirical) data point, that is $\hat{\mu}(B_{\mathbf{i}}^J) \neq 0$
	- $\frac{\lambda(B_{\mathbf{i}})}{\lambda(B_{\mathbf{i}}^J)}$ otherwise.
- Simulate uniformly from $B_{\mathbf{i}}^{-J}$ 

This algorithm simulates first the known part of the model (dimensions in $J$), and then, conditionally, the checkerboard part, ensuring that the known copula is respected. The downside of this behavior is that the checkerboard part may have points outside standard checkerboard boxes, making this part of the copula less sparse than a true checkerboard. But is does become sparser as soon as the data fits the known margins. On the other hand, this algorithm allows for a lot of flexibility, mainly in the following points : 


- The "grid" given by $\mathcal{B}_\mathcal{I}$ can be taken more arbitrarily than $m^d$ boxes of same volume, as soon as it's a partition of $\mathbb{I}^d$.
- The known copula is *not* restricted at all and can be choose among all p-variates copulas. 
- The estimation of the checkerboard part can be turn into a more flexible *patchwork* construction, by changing the independence copula for an other one inside the boxes. See Durante2013,Durante2015,Durante2015a

We are now going to define properly the measure associated to this simulation procedure, a.k.a the empirical checkerboard copula with known margins. Let $\nu$ be this measure and let $\mathbf{U}$ be a random vector drawn from it. Then $\forall \mathbf{x} \in \mathbb{I}^d$, following the above procedure, we have :  

$$
	\begin{split}
	\nu([0,\mathbf{x}]) &= \mathbb{P}(\mathbf{U}\le \mathbf{x}) \\
			   &= \sum\limits_{{\mathbf{i}} \in \mathcal{I}} \mathbb{P}(\mathbf{U} \in B_{\mathbf{i}}, \mathbf{U} \le \mathbf{x}) \\
			   &= \sum\limits_{{\mathbf{i}} \in \mathcal{I}} \mathbb{P}(\mathbf{U}^{-J} \in B_{\mathbf{i}}^{-J} \cap [0,\mathbf{x}^{-J}] | \mathbf{U}^{J} \in B_{\mathbf{i}}^{J} \cap [0,\mathbf{x}^{J}]) \mathbb{P}(\mathbf{U}^{J} \in B_{\mathbf{i}}^{J} \cap [0,\mathbf{x}^{J}])
	\end{split}
$$			
			
While the unconditional term is easy to handle since it's the measure associated with the known copula, $\mathbb{P}(\mathbf{U}^{J} \in B_{\mathbf{i}}^{J} \cap [0,\mathbf{x}^{J}]) = \mu^J(B_{\mathbf{i}}^{J} \cap [0,\mathbf{x}^{J}])$, the conditional term can be treated according to the algorithm : it will be $\frac{\lambda(B_{\mathbf{i}}^{-J} \cap [0,\mathbf{x}^{-J}])}{\lambda(B_{\mathbf{i}}^{-J})}$ inside a box choosed with probability conditional on $\hat{\mu}(B_{\mathbf{i}}^J) \neq 0$. We finaly get the following definition :



**The empirical checkerboard copula with parameter $m$ and with set of known margins $J$ following the measure $\mu^J$ is the copula corresponding to the measure $\nu$ given by : **

$$
\nu([0,x]) = \sum\limits_{{\mathbf{i}} \in \mathcal{I}} \mu^J(B_{\mathbf{i}}^{J} \cap [0,\mathbf{x}^{J}]) \frac{\lambda(B_{\mathbf{i}}^{-J} \cap [0,\mathbf{x}^{-J}])}{\lambda(B_{\mathbf{i}}^{-J})} \left[ \frac{\hat{\mu}(B_{\mathbf{i}})}{\hat{\mu}(B_{\mathbf{i}}^J)} \mathbb{1}_{\hat{\mu}(B_{\mathbf{i}}^J) \neq 0} + \frac{\lambda(B_{\mathbf{i}})}{\lambda(B_{\mathbf{i}}^J)} \mathbb{1}_{\hat{\mu}(B_{\mathbf{i}}^J) = 0}   \right]
$$



The next section will discuss the current implementation of this copula;

## Current implementation 

The package implemnts the ECBC with known margins through the `cbkmCopula` class. The constructor of the class takes several arguments : 

- `x`, representing the pseudo_data.
- `m=nrow(x)`, repesenting the checkerboard parameter
- `pseudo=FALSE`, representing wether or not the pseudo_data is already given in a pseudo_observation form.
- `margins_numbers=NULL`, representing the margins index that are associated to the known copula, formerly noted $J$
- `known_cop=NULL`, representing the known copula to be applied to thoose margins : a copula object of right imension. 

for exemple, let's work on some simulated datas, e.g a 4-dimension clayton copula simulated via the `copula` package.

```{r,fig.cap="Pairs-plot of the simulated dataset"}
  true_copula <- onacopulaL(
    family = "Clayton",
    nacList = list(iTau(getAcop("Clayton"), 0.6), 1:4)
  )
  dataset <- rCopula(50,true_copula) 
  colnames(dataset) <- c("u","v","w","x")
  pairs(dataset,lower.panel=NULL)

```


And let's suppose that the second bivariate margin is known, with a (right) clayton copula with $\tau = 0.6$ : 
```{r}
  known_margins <- c(2,3)
  known_clayton <- onacopulaL(
    family = "Clayton",
    nacList = list(iTau(getAcop("Clayton"), 0.6), 1:2)
  )
```

Then we can construct the ECBC with this known margin : 

```{r}
  cop <- cbkmCopula(x = dataset,m = 5,pseudo = TRUE,margins_numbers = known_margins,known_cop = known_clayton)
  cop
```

We can then simulate from it : 

```{r, fig.cap="Pairs-plot of the original data (red) and simulated data from the true model (black)"}
  simu <- rCopula(500,cop)
  pairs(rbind(simu,dataset),col=c(rep("black",nrow(simu)),rep("red",nrow(dataset))),gap=0,lower.panel = NULL)
```



You can see that the known-margin was respected. 

What now if the known margins is clearly missspecified ? 
```{r, fig.cap="Pairs-plot of the original data (red) and simulated data from a wrong model (black)"}
  wrong_clayton <- onacopulaL(
    family = "Clayton",
    nacList = list(iTau(getAcop("Clayton"), 0.1), 1:2)
  )
  cop <- cbkmCopula(x = dataset,m = 5,pseudo = TRUE,margins_numbers = known_margins,known_cop = wrong_clayton)

  simu <- rCopula(500,cop)
  pairs(rbind(simu,dataset),col=c(rep("black",nrow(simu)),rep("red",nrow(dataset))),gap=0,lower.panel = NULL)
```


The conditioning mecanisme did not do well for other margins too. 

What now if the 2 parts are clearly independant ? 

```{r, fig.cap="Pairs-plot of the original data with independance"}

  true_copula1 <- onacopulaL(
    family = "Clayton",
    nacList = list(iTau(getAcop("Clayton"), 0.8), 1:2)
  )
  true_copula2 <- onacopulaL(
    family = "Clayton",
    nacList = list(iTau(getAcop("Clayton"), 0.8), 1:2)
  )
  
  dataset <- cbind(rCopula(100,true_copula1),rCopula(100,true_copula2))
  colnames(dataset) <- c("u","v","w","x")
  pairs(dataset,lower.panel=NULL)

```
```{r, fig.cap="Pairs-plot of the original data (red) and simulated data from a wrong model (black) -- Independance case"}
  cop <- cbkmCopula(x = dataset,m = 10,pseudo = TRUE,margins_numbers = c(1,2),known_cop = wrong_clayton)

  simu <- rCopula(500,cop)
  pairs(rbind(simu,dataset),col=c(rep("black",nrow(simu)),rep("red",nrow(dataset))),gap=0,lower.panel = NULL)
```

We can see that the wrong specification of the clayton copula for the 2 firsts margins (u,v) impacted less the checkerboard construction for the other 2 margins (w,x), since the two parts are (truly) independant.





